---
title: "Germination drivers preliminary analysis"
author: "Clara Espinosa"
date: "2024-01-29"
output: html_document
---
# Large header
## Not so large header
### Even smaller header
**This is bold**, *This is italic*, ***Both bold and italic***
[text that will appear](add link)  
Showing images  
<center> where to put the image  
![description of an image](link of the image)  
</center> stop taking this into account  
{height=300px} to format images  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
library(tidyverse);library(readxl);library(ggplot2);library(dplyr)
library(viridis);library(rstatix);library(binom)
# echo=FALSE, code is evaluated but not run
```

# Cleaning and preliminary exploration of the data   

## Load raw data  

### Germination raw data  
```{r, echo= FALSE}
read.csv("data/raw_data.csv", sep = ",") %>%
  convert_as_factor(species, code, treatment, temp) -> raw_df
str(raw_df)
library(kableExtra) # fuuthr and detailed formatting of tables! explore options
unique(raw_df$species)
knitr::kable(head(raw_df), caption ="Raw germination data")%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive", position = "center"))

```

### Species information data
```{r}
read.csv("data/species.csv", sep=",") %>%
  dplyr::select(species, code, temperature_regime, family, community, habitat)  %>%
  convert_as_factor(species, code, temperature_regime, family, community, habitat)  -> species
str(species)
head(species)
unique(species$species)
```

### List of species with 0 germination, consider removing from further analysis: 


  
```{r eval=FALSE}
  filter (!species == "Euphrasia salisburgensis")%>%
  filter (!species == "Gentiana verna")%>%
  filter (!species == "Gentianella campestris")%>%
  filter (!species == "Kobresia myosuroides")%>%
  filter (!species == "Salix breviserrata")%>% # also with less than 25% of seeds viables
  filter (!species == "Sedum album")%>%
  filter (!species == "Sedum atratum")%>%
  filter (!species == "Solidago virgaurea")
```

## Basic calculations  
### Viable seed calculation per petri dish
```{r}
read.csv("data/raw_data.csv", sep = ",") %>%
  convert_as_factor(species, code, treatment, temp) %>%
  dplyr::select (species, code, treatment, petri, initial, empty, fungus) %>%
  group_by (species, code, treatment, petri) %>%
  mutate (viable = initial -(empty + fungus)) %>%
  dplyr::select (species, code, treatment, petri, viable) -> viables_petri
unique(viables_petri$species)
head(viables_petri)
```

### Viables calculation per community
```{r}
read.csv("data/raw_data.csv", sep = ",") %>%
    convert_as_factor(species, code, treatment, temp) %>%
    dplyr::select (species, code, treatment, petri, initial, empty, fungus) %>%
    group_by (species, code, treatment, petri) %>%
    mutate (viable = initial -(empty + fungus)) %>%
    dplyr::select (species, code, treatment, petri, viable)%>%
    merge(species)%>%
  filter (!species == "Euphrasia salisburgensis")%>%
  filter (!species == "Gentiana verna")%>%
  filter (!species == "Gentianela campestris")%>%
  filter (!species == "Kobresia myosuroides")%>%
  filter (!species == "Salix breviserrata")%>%
  filter (!species == "Sedum album")%>%
  filter (!species == "Sedum atratum")%>%
  filter (!species == "Solidago virgaurea")%>%
    group_by (community, treatment)%>%
    summarise(viable= sum(viable))-> viables_community
viables_community
```

### Viables calculation per species and treatment
```{r}
read.csv("data/raw_data.csv", sep = ",") %>%
    convert_as_factor(species, code, treatment, temp) %>%
    dplyr::select (species, code, treatment, petri, initial, empty, fungus) %>%
    group_by (species, code, treatment, petri) %>%
    mutate (viable = initial -(empty + fungus)) %>%
    dplyr::select (species, code, treatment, petri, viable)%>%
    group_by (species,code, treatment)%>%
    summarise(viable= sum(viable))-> viables_sp
unique(viables_sp$species)
head(viables_sp)
```
 
### Germination during cold stratification period
```{r}
read.csv("data/raw_data.csv", sep = ",") %>%
  convert_as_factor(species, code, treatment, temp) %>%
  merge(viables_petri) %>%
  dplyr::select (species, code, petri, viable, initial, cold_strat) %>% ## 
  na.omit() %>% #remove alternate dark treatment rows
  group_by (species, code,petri) %>%
  dplyr::summarize(finalgerm = sum(cold_strat),
            viable = sum(viable)) %>% 
  mutate (finalgerm = replace_na(finalgerm , 0)) %>% # salix and solidago have some petridish with all fungus
  mutate (treatment = "E_cold_stratification") %>%
  mutate(treatment = as.factor(treatment))-> cold_strat 
unique(cold_strat$species)
head(cold_strat)
```

#### Visualization of germination during cold stratification pre-treatment with binomial errors
```{r, echo = F, fig.height=8, fig.width=10, fig.cap="Germination during cold stratification", fig.align="center", dpi=300}
cold_strat %>%
  mutate (binom.confint(finalgerm, viable, methods = "wilson"))%>%
  dplyr::select (species, code,  treatment, finalgerm, viable, mean, upper, lower)%>%
  group_by (species)%>%
  dplyr::summarize(germpro = mean(mean), 
            upper = mean(upper),
            lower = mean (lower)) %>% 
  ggplot(aes(x= species, y= germpro, fill = species)) +
  geom_bar (stat = "identity") +
  geom_errorbar( aes(species, germpro, ymin = lower, ymax = upper), width = 0.5, linewidth = 0.5) +
  coord_flip() + 
  labs (title = "Germination during cold stratification", x = "Species", y = "Germination proportion")+
  theme_minimal (base_size = 16) +
  theme (plot.title = element_text ( size =20), #hjust = 0.5,
    legend.position = "none")

```

### Separate datsets per community with germination during cold stratification period
```{r}
cold_strat %>%
  group_by(species, code, treatment)%>%
  dplyr::summarize(finalgerm = sum(finalgerm),
                   viable = sum(viable))%>%
  merge(species)%>%
  filter(community=="Mediterranean")%>%
  dplyr::select(code, species, treatment, viable, finalgerm)->cold_strat_M; cold_strat_M

cold_strat %>%
  group_by(species, code, treatment)%>%
  dplyr::summarize(finalgerm = sum(finalgerm),
                   viable = sum(viable))%>%
  merge(species)%>%
  filter(community=="Temperate")%>%
  dplyr::select(code, species, treatment, viable, finalgerm)->cold_strat_T;cold_strat_T
```

### Final germination percentage per treatment discounting cold stratification germ in darkness treatment 
```{r}
read.csv("data/raw_data.csv", sep = ",") %>%
  convert_as_factor(species, code, treatment, temp) %>%
  dplyr::select (species, code, treatment, petri, initial, D7, D14, D21, D28, D35, D42) %>%
  gather ("scores", "germ", D7:D42) %>%
  mutate (germ = replace_na(germ, 0)) %>%
  group_by (species, code, treatment, petri)%>%
  dplyr::summarize (finalgerm = sum(germ)) %>% # 
  dplyr::select(species, code, treatment, petri, finalgerm)%>% 
  merge(viables_petri, by= c("species", "code", "treatment", "petri")) %>%
  rbind(cold_strat)%>%
  group_by (species, code, treatment) %>%
  dplyr::summarize(finalgerm = sum (finalgerm)) %>% 
  spread(treatment, finalgerm)%>%
  mutate (B_alternate_dark = B_alternate_dark - E_cold_stratification)  %>% # substract cold stratification
  mutate (B_alternate_dark = ifelse(B_alternate_dark < 0, 0, B_alternate_dark)) %>% # convert negative values to 0 germination
  gather ("treatment", "finalgerm", A_alternate_light:E_cold_stratification) %>%
  filter(!treatment=="E_cold_stratification")%>%
  merge(viables_sp, by= c("code", "species", "treatment"))-> finalgerm;finalgerm
str(finalgerm)
head(finalgerm)
```
### Graph comparing final germiantion between treatment at community level
```{r echo=FALSE}
#echo=FALSE code not showed
finalgerm %>%
  merge(species)%>%
  filter (!species == "Euphrasia salisburgensis")%>% # species with 0 germination
  filter (!species == "Gentiana verna")%>%
  filter (!species == "Gentianella campestris")%>%
  filter (!species == "Kobresia myosuroides")%>%
  filter (!species == "Salix breviserrata")%>%
  filter (!species == "Sedum album")%>%
  filter (!species == "Sedum atratum")%>%
  filter (!species == "Solidago virgaurea") %>%
  group_by (community, species, treatment) %>%
  na.omit() %>%
  group_by (treatment, community) %>% #, temperature_regime 
  summarise(finalgerm=sum(finalgerm),
            viable = sum(viable))%>%
  mutate (binom.confint(finalgerm, viable, methods = "wilson"))%>%
  ggplot()+
  geom_bar(aes(x= treatment, y= mean, fill= treatment), stat = "identity") +
  geom_errorbar(aes(x= treatment, y=mean, ymin=lower, ymax=upper), color = "black", size=1, width = 0.4)+
  facet_grid(~community) +
  #ylim (0,1)+
  scale_fill_viridis_d(name = "Treatments",labels = c("Control", "Darkness", "Water stress", "Constant Temperature", "Cold stratification")) +
  labs (title = "Mean Germination per Treatment", x = "Treatments", y="Germination proportion")+
  theme_classic (base_size = 16) + #theme_minimal for all species for mean treatment
  theme (plot.title = element_text ( size = 24), #hjust = 0.5,
         strip.text = element_text (size =20),
         axis.text.x= element_blank(), 
         legend.position = "right")
```

#### Cumulative germination graph for each treatment at community level
```{r echo=FALSE}
raw_df %>%
  mutate ( D0 = 0) %>%
  dplyr::select (species, treatment, initial, D0, D7, D14, D21, D28, D35, D42) %>%
  gather ("scores", "germ", D0:D42) %>%
  filter (!(treatment == "alternate_dark")) %>%
  mutate (germ = replace_na(germ, 0)) %>%
  mutate (scores = factor(scores, levels = c("D0","D7", "D14", "D21", "D28","D35", "D42"))) %>%
  arrange(species, treatment, scores)%>% 
  merge (species) %>%
  filter (!species == "Euphrasia salisburgensis")%>%
  filter (!species == "Gentiana verna")%>%
  filter (!species == "Gentianela campestris")%>%
  filter (!species == "Kobresia myosuroides")%>%
  filter (!species == "Salix breviserrata")%>%
  filter (!species == "Sedum album")%>%
  filter (!species == "Sedum atratum")%>%
  filter (!species == "Solidago virgaurea")%>%
  group_by (community, treatment, scores)%>%
  summarise (germinated = sum(germ))%>%
  mutate (cum_germ = cumsum(germinated))  %>% #
  merge(viables_community) %>%
  mutate(germpro = (cum_germ/viable), # 
         germpro = round (germpro, digit =2)) %>%
  filter(!treatment == "B_alternate_dark")%>%
  ggplot(aes(x= scores, y= germpro, color= treatment, group = treatment))+
  geom_line(linewidth = 1.5) +
  facet_grid(~community) +
  #ylim (0,1)+
  scale_color_manual (name= "Treatment", values = c ("C_alternate_WP"= "#fde725", "A_alternate_light" ="#21918c", "D_constant_light" = "#5ec962")) +
  #scale_color_viridis_d() +
  labs (title = "Cumulative Germination")+
  xlab ("Germination scores") +
  ylab("Germination proportion") +
  theme_classic (base_size = 16) + #theme_minimal for all species for mean treatment
  theme (plot.title = element_text ( size = 32), #hjust = 0.5,
         legend.position = "right")
```

# Species preferences 
```{r echo=F}
library (tidyverse); library (dplyr);library(rstatix); library(ggrepel)
library (ggpubr);library (vegan)
```
Species preferences calculated from ibuttons data and inventories the original scripts and data in Omaña80 and Picos Github repository (respectively) script name: indices and glms 
climatic indices used:   
FROM ibuttons data: in picos from 02/10/2018 to 07/08/2019; in villabandin from 12/07/2021 to 29/05/2022  
FDD: sum of degrees (daily mean) when daily Tmean is below 0 ºC (in absolute values)  
GDD: sum of degrees (daily mean) when daily Tmean is above 5 ºC   
Snowdays = sum of days with snow cover, calculated as days when Tmax <0.5 and Tmin > -0.5 ºC  
These variables (altogether with bio1, bio2 and bio7) calculated mean x day, then month and finally year (whole datset)

From inventory data: 80 plots per each community, at each plot species coverage estimated in %
To calculate species preferences several filters applied:
    # consider only plots where each species has at least 10% coverage
    # climatic variables weighted per coverage 
